---
title: "Phagocytosis"
author: "Shakir Suleimanov"
date: "2023-06-02"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: true
    toc_depth: 3
    theme: cerulean
    keep_md: true
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE, results = FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=TRUE, warning = FALSE, message = FALSE, results = FALSE, error=FALSE)

library(tidyverse)
library(readxl)
library(openxlsx)
library(ggfortify) 
library(pheatmap)
library(extrafont)
library(tibble)
library(viridis)
library(hrbrthemes)
library(multcomp)
```
##Normalized data for flow cytometry - percent of phagocytosis

```{r}
flow_cytometry_percents_10_1 <- read_excel("Database.xlsx", sheet = "%pos_flow_cytometry")

flow_cytometry_percents_scaling_10_1_for_clusters <- flow_cytometry_percents_10_1 %>%
  filter(!Group == "Experiment4") %>%
  dplyr::select(is.numeric) %>%
  t() %>%
  scale()
  
flow_cytometry_percents_scaling_10_1 <- flow_cytometry_percents_10_1 %>%
  filter(!Group == "Experiment4") %>%
  dplyr::select(is.numeric) %>%
  t() %>%
  scale() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rowwise() %>%
  mutate(average_scale = mean(c_across(contains("V")))) %>%
  mutate(rowname = as.factor(rowname))

colnames(flow_cytometry_percents_scaling_10_1) <- c("Group", "Experiment1", "Experiment2", "Experiment3", "Experiment4", "Average")

flow_cytometry_percents_scaling_10_1_ANOVA <- flow_cytometry_percents_scaling_10_1 %>%
  pivot_longer(!Group & !Average)

tiff("Figures/boxplot_flow_cytometry_percents_norm.tiff", width = 6, height = 4, units = 'in', res = 360, compression = 'none')
phm_fc_percents <- pheatmap(flow_cytometry_percents_scaling_10_1_for_clusters, cluster_col = TRUE, show_colnames = TRUE, annotation_names_row = T, fontsize_col = 8, cutree_rows = 2)
dev.off()
```

```{r}
ANOVA_perc_flow_cytometry <- aov(value ~ Group, data = flow_cytometry_percents_scaling_10_1_ANOVA) %>% glht(linfct = mcp(Group = "Tukey")) %>% summary()

ANOVA_perc_flow_cytometry

aov(value ~ Group, data = flow_cytometry_percents_scaling_10_1_ANOVA) %>% summary()
```

#Bubbleplot for percents flow cytometry
```{r}
plot_for_bubblegum_fc_10_1_scaled <- flow_cytometry_percents_scaling_10_1 %>%
  pivot_longer(!Group) %>%
  filter(name == "Average")
  
bubblegum_plot_perc_flow_cytometry_10_1 <- ggplot(plot_for_bubblegum_fc_10_1_scaled, aes(x = Group, y = name, fill = value, size = value)) +
  geom_point(alpha = 0.8, shape = 21, stroke = 0) +
  #geom_hline(yintercept = seq(.5, 4.5, 1), size = .2) +
  scale_x_discrete(position = "bottom") +
  scale_radius(range = c(5, 40), breaks = c(-1, 0, 1), labels = c(-1, 0, 1), limits = c(-2, 2)) +
  scale_fill_gradient(low = "orange", high = "blue", breaks = c(-2, 0, 2), labels = c("Low", "Middle", "High"), limits = c(-2, 2)) +
  theme_ipsum() +
  theme(legend.position = "bottom", 
        panel.grid.major = element_blank(),
        legend.text = element_text(size = 14, family="Arial"),
        legend.title = element_text(size = 14, family="Arial"),
        axis.text = element_text(size = 14, family = "Arial"), axis.text.x=element_text(angle = 45, hjust = 0.9)) +
  guides(size = guide_legend(override.aes = list(fill = "grey", alpha = 0.5, color = "black", stroke = .25), label.position = "bottom",title.position = "left", order = 1), fill = guide_colorbar(ticks.colour = NA, title.position = "top", order = 2)) +
  labs(size = "z-scores", fill = "Level of phagocytosis:", x = NULL, y = NULL)

bubblegum_plot_perc_flow_cytometry_10_1 

ggsave("Figures/bg_plot_fc_norm_total_perc.tiff")
ggsave("Figures/bg_plot_fc_norm_total_perc.jpeg")
```

## Supplementary materials. percent flow cytometry with boxplots 

```{r}
data_M1_MCSF_GM_CSF <- read_excel("Database.xlsx", sheet = "%pos_flow_cytometry")

perc_pos_flow_cytometry_for_boxplot <- data_M1_MCSF_GM_CSF %>%
  filter(!Group == "Experiment4") %>%
  dplyr::select(-Date & -Ratio) %>%
  pivot_longer(!Group) %>%
  mutate(name = as.factor(name))
  
boxplot_perc_pos_flow_cytometry <- ggplot(perc_pos_flow_cytometry_for_boxplot, aes(x = name, y = value, fill = name)) +
  geom_boxplot(alpha = 0.6) +
  geom_point(size = 3) +
  scale_fill_manual(breaks = perc_pos_flow_cytometry_for_boxplot$name,
                    values = c("purple", "blue", "orange", "red")) +
  theme(axis.title.x=element_blank(), axis.text.y=element_text(angle=0,hjust=0.5,vjust=0.4, size = 18),
axis.text.x=element_text(angle=0,hjust=0.5,vjust=0.4, size = 18),
        legend.text = element_text(size = 18, family="Arial"),
        legend.title = element_text(size = 18, family="Arial"))

boxplot_perc_pos_flow_cytometry

model_percent_flow_cytometry <- aov(value ~ name, data = perc_pos_flow_cytometry_for_boxplot) %>% glht(linfct = mcp(name = "Tukey")) %>% summary()

model_percent_flow_cytometry

ggsave("Figures/boxplot_perc_pos_flow_cytometry.tiff")
ggsave("Figures/boxplot_perc_pos_flow_cytometry.jpeg")
```


##Normalized data for flow cytometry - MFI

```{r}
flow_cytometry_MFI_10_1 <- read_excel("Database.xlsx", sheet = "MFI_flow_cytometry")

flow_cytometry_MFI_scaling_10_1_for_clusters <- flow_cytometry_MFI_10_1 %>%
  filter(!Group == "Experiment4") %>%
  dplyr::select(is.numeric) %>%
  t() %>%
  scale()
  
flow_cytometry_MFI_scaling_10_1 <- flow_cytometry_MFI_10_1 %>%
  filter(!Group == "Experiment4") %>%
  dplyr::select(is.numeric) %>%
  t() %>%
  scale() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rowwise() %>%
  mutate(average_scale = mean(c_across(contains("V")))) %>%
  mutate(rowname = as.factor(rowname))

colnames(flow_cytometry_MFI_scaling_10_1) <- c("Group", "Experiment1", "Experiment2", "Experiment3", "Experiment4", "Average")

flow_cytometry_MFI_scaling_10_1_ANOVA <- flow_cytometry_MFI_scaling_10_1 %>%
  pivot_longer(!Group & !Average)


tiff("Figures/phm_MFI_flow_cytometry.tiff", width = 6, height = 4, units = 'in', res = 360, compression = 'none')
phm_fc_MFI <- pheatmap(flow_cytometry_MFI_scaling_10_1_for_clusters, cluster_col = TRUE, show_colnames = TRUE, annotation_names_row = T, fontsize_col = 8, cutree_rows = 2)
dev.off()
```

```{r}
ANOVA_MFI_flow_cytometry <- aov(value ~ Group, data = flow_cytometry_MFI_scaling_10_1_ANOVA) %>% glht(linfct = mcp(Group = "Tukey")) %>% summary()

ANOVA_MFI_flow_cytometry

aov(value ~ Group, data = flow_cytometry_MFI_scaling_10_1_ANOVA)%>%summary()
```



```{r}
plot_for_bubblegum_fc_MFI_10_1_scaled <- flow_cytometry_MFI_scaling_10_1 %>%
  pivot_longer(!Group) %>%
  filter(name == "Average")
  
bubblegum_plot_MFI_flow_cytometry_10_1 <- ggplot(plot_for_bubblegum_fc_MFI_10_1_scaled, aes(x = Group, y = name, fill = value, size = value)) +
  geom_point(alpha = 0.8, shape = 21, stroke = 0) +
  #geom_hline(yintercept = seq(.5, 4.5, 1), size = .2) +
  scale_x_discrete(position = "bottom") +
  scale_radius(range = c(5, 40), breaks = c(-1, 0, 1), labels = c(-1, 0, 1), limits = c(-2, 2)) +
  scale_fill_gradient(low = "orange", high = "blue", breaks = c(-2, 0, 2), labels = c("Low", "Middle", "High"), limits = c(-2, 2)) +
  theme_ipsum() +
  theme(legend.position = "bottom", 
        panel.grid.major = element_blank(),
        legend.text = element_text(size = 14, family="Arial"),
        legend.title = element_text(size = 14, family="Arial"),
        axis.text = element_text(size = 14, family = "Arial"), axis.text.x=element_text(angle = 45, hjust = 0.9)) +
  guides(size = guide_legend(override.aes = list(fill = "grey", alpha = 0.5, color = "black", stroke = .25), label.position = "bottom",title.position = "left", order = 1), fill = guide_colorbar(ticks.colour = NA, title.position = "top", order = 2)) +
  labs(size = "z-scores", fill = "Level of phagocytosis:", x = NULL, y = NULL)

bubblegum_plot_MFI_flow_cytometry_10_1 

ggsave("Figures/bg_plot_flow_MFI_norm_total.tiff")
ggsave("Figures/bg_plot_flow_MFI_norm_total.jpeg")
```


## Supplementary materials. MFI flow cytometry with boxplots 

```{r}
data_M1_MCSF_GM_CSF_MFI <- read_excel("Database.xlsx", sheet = "MFI_flow_cytometry")

MFI_flow_cytometry_for_boxplot <- data_M1_MCSF_GM_CSF_MFI %>%
  filter(!Group == "Experiment4") %>%
  dplyr::select(-Date & -Ratio) %>%
  pivot_longer(!Group) %>%
  mutate(name = as.factor(name))
  
boxplot_MFI_flow_cytometry <- ggplot(MFI_flow_cytometry_for_boxplot, aes(x = name, y = value, fill = name)) +
  geom_boxplot(alpha = 0.6) +
  geom_point(size = 3) +
  scale_fill_manual(breaks = perc_pos_flow_cytometry_for_boxplot$name,
                    values = c("purple", "blue", "orange", "red")) +
  theme(axis.title.x=element_blank(), axis.text.y=element_text(angle=0,hjust=0.5,vjust=0.4, size = 18),
axis.text.x=element_text(angle=0,hjust=0.5,vjust=0.4, size = 18),
        legend.text = element_text(size = 18, family="Arial"),
        legend.title = element_text(size = 18, family="Arial"))

boxplot_MFI_flow_cytometry

model_MFI_flow_cytometry <- aov(value ~ name, data = MFI_flow_cytometry_for_boxplot) %>% glht(linfct = mcp(name = "Tukey")) %>% summary()

model_MFI_flow_cytometry

ggsave("Figures/boxplot_MFI_flow_cytometry.tiff")
ggsave("Figures/boxplot_MFI_flow_cytometry.jpeg")
```


##Normalized data for fluorescence microscopy - percent

```{r}
fluor_microscopy_perc <- read_excel("Database.xlsx", sheet = "%pos_f_microscopy")

fluor_microscopy_perc_for_clusters <- fluor_microscopy_perc %>%
  dplyr::select(is.numeric) %>%
  t() %>%
  scale()
  
fluor_microscopy_perc_bgplot <- fluor_microscopy_perc %>%
  dplyr::select(is.numeric) %>%
  t() %>%
  scale() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rowwise() %>%
  mutate(average_scale = mean(c_across(contains("V")))) %>%
  mutate(rowname = as.factor(rowname))

colnames(fluor_microscopy_perc_bgplot) <- c("Group", "Experiment1", "Experiment2", "Experiment3", "Experiment4", "Average")

tiff("Figures/phm_percent_fluorescence_microscopy.tiff", width = 6, height = 4, units = 'in', res = 360, compression = 'none')
pheatmap(fluor_microscopy_perc_for_clusters, cluster_col = TRUE, show_colnames = TRUE, annotation_names_row = T, fontsize_col = 8, cutree_rows = 2)
dev.off()
```


```{r }
fluor_microscopy_perc_bgplot_ANOVA <- fluor_microscopy_perc_bgplot %>%
  pivot_longer(!Group & !Average)

ANOVA_percent_fluorescence_microscopy <- aov(value ~ Group, data = fluor_microscopy_perc_bgplot_ANOVA) %>% glht(linfct = mcp(Group = "Tukey")) %>% summary()

ANOVA_percent_fluorescence_microscopy
```


```{r}
fluor_microscopy_perc_bgplot_scaled <- fluor_microscopy_perc_bgplot %>%
  pivot_longer(!Group) %>%
  filter(name == "Average")
  
bubblegum_plot_perc_f_microscopy <- ggplot(fluor_microscopy_perc_bgplot_scaled, aes(x = Group, y = name, fill = value, size = value)) +
  geom_point(alpha = 0.8, shape = 21, stroke = 0) +
  #geom_hline(yintercept = seq(.5, 4.5, 1), size = .2) +
  scale_x_discrete(position = "bottom") +
  scale_radius(range = c(5, 40), breaks = c(-1, 0, 1), labels = c(-1, 0, 1), limits = c(-2, 2)) +
  scale_fill_gradient(low = "orange", high = "blue", breaks = c(-2, 0, 2), labels = c("Low", "Middle", "High"), limits = c(-2, 2)) +
  theme_ipsum() +
  theme(legend.position = "bottom", 
        panel.grid.major = element_blank(),
        legend.text = element_text(size = 8, family="Optima"),
        legend.title = element_text(size = 8, family="Optima"),
        axis.text = element_text(size = 10, family = "Optima"), axis.text.x=element_text(angle = 45, hjust = 0.9)) +
  guides(size = guide_legend(override.aes = list(fill = "grey", alpha = 0.5, color = "black", stroke = .25), label.position = "bottom",title.position = "right", order = 1), fill = guide_colorbar(ticks.colour = NA, title.position = "top", order = 2)) +
  labs(size = "z-scores", fill = "Level of phagocytosis:", x = NULL, y = NULL)

bubblegum_plot_perc_f_microscopy

ggsave("Figures/bg_plot_percent_fluorescence_microscopy.tiff")
ggsave("Figures/bg_plot_percent_fluorescence_microscopy.jpeg")
```


## Supplementary materials. percent fluorescence microscopy with boxplots 

```{r}
data_M1_MCSF_GM_CSF_percent_microscopy <- read_excel("Database.xlsx", sheet = "%pos_f_microscopy")

percent_f_microscopy_for_boxplot <- data_M1_MCSF_GM_CSF_percent_microscopy %>%
  dplyr::select(-Date & -Ratio) %>%
  pivot_longer(!Group) %>%
  mutate(name = as.factor(name))
  
boxplot_percent_f_microscopy <- ggplot(percent_f_microscopy_for_boxplot, aes(x = name, y = value, fill = name)) +
  geom_boxplot(alpha = 0.6) +
  geom_point(size = 3) +
  scale_fill_manual(breaks = percent_f_microscopy_for_boxplot$name,
                    values = c("purple", "blue", "orange", "red")) +
  theme(axis.title.x=element_blank(), axis.text.y=element_text(angle=0,hjust=0.5,vjust=0.4, size = 18),
axis.text.x=element_text(angle=0,hjust=0.5,vjust=0.4, size = 18),
        legend.text = element_text(size = 18, family="Arial"),
        legend.title = element_text(size = 18, family="Arial"))

boxplot_percent_f_microscopy

model_percent_f_microscopy <- aov(value ~ name, data = percent_f_microscopy_for_boxplot) %>% glht(linfct = mcp(name = "Tukey")) %>% summary()

model_percent_f_microscopy

ggsave("Figures/boxplot_percent_f_microscopy.tiff")
ggsave("Figures/boxplot_percent_f_microscopy.jpeg")
```


## Cytometry*Microscopy correlation

```{r}
cor_data <- read_excel("Database.xlsx", sheet = "cor_fc_mic")

library(ggcorrplot)

cor_table <- cor_data %>%
  pivot_wider(., names_from = exp, values_from = c(fc, mic)) %>%
  dplyr::select(is.numeric)

cor_table_2 <- cor_data %>%
  dplyr::select(is.numeric)

p.mat = cor_pmat(cor_table)

cor_plot_fc_mic <- ggcorrplot(cor(cor_table), hc.order = FALSE, method = "circle", p.mat = p.mat,
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   colors = c("#6D9EC1","white", "#E46726")) +
  theme(axis.text.x=element_text(angle = 45))

cor_plot_fc_mic

ggsave("Figures/cor_table.tiff")


cor_plot_fc_mic_2 <- ggcorrplot(cor(cor_table_2), hc.order = TRUE, method = "circle", p.mat = p.mat,
   outline.col = "white",
   ggtheme = ggplot2::theme_gray,
   colors = c("#6D9EC1","white", "#E46726")) +
  theme(axis.text.x=element_text(angle = 45))

cor_plot_fc_mic_2

ggplot(cor_data) +
  geom_point(aes(x = fc, y = mic, color = group, size = exp), alpha = 0.7) +
  geom_smooth(aes(x = fc, y = mic), method = "lm") +
  scale_fill_manual(breaks = cor_data$group,
                    values = c("purple", "blue", "orange", "red")) +
  theme(axis.title.x=element_blank(), axis.text.y=element_text(angle=0,hjust=0.5,vjust=0.4, size = 16),
axis.text.x=element_text(angle=0,hjust=0.5,vjust=0.4, size = 16),
        legend.text = element_text(size = 16, family="Arial"),
        legend.title = element_text(size = 16, family="Arial"))

ggsave("Figures/corplot.tiff")

cor_coeff <- cor(cor_data$fc, cor_data$mic, method = "pearson")

cor_coeff
```

